\bf A Hierarchical Bayesian Approach for Modeling Infant-Mortality and Wearout Failure Modes

### 2017-07-17 paper/paper.tex
Merged my edits with Eric's from earlier in the day.

ADD PIECE ABOUT TRANSFORMATIONS

For estimation, we modeled the data on the log scale, using the Smallest Extreme Value (SEV) distribution \cite{meeker}.  If $\text{log}(Y) \sim \text{SEV}(\mu, \sigma)$ then $Y \sim \text{Weibull}(\beta = 1/\sigma, \eta = \exp(\mu))$.  This decision was chosen after encountering numerical issues when performing estimation using the standard Weibull parametrization.  In addition, we transformed the parameters to the lower log quantile, $\log(t_{p_j})$ (where $p_{1} = 0.5$, and $p_2 = 0.2$), and log scale, $\log(\sigma_j)$.  This parameterization was selected as the majority of failures occur at the lower tail of the distribution.  \\

@@

\subsection{Censoring and Truncation}
Right censoring is common in reliability data.  Due to time or budget constraints it is rare all units are observed until failure. If a unit has not yet failed when the data is analyzed it is considered right censored.  In other words, right censoring puts a lower bound on the failure time.  Products with long expected lifetimes, such as hard drives, are heavily right censored.  Hard drives are built to last for decades under normal use conditions.  Even when running continuously, hard drives can exhibit failure rates of less than two percent per year \cite{backblaze}.  \\

While right censoring imposes an upper bound on an observation, truncation means the existence of an observation is unknown if it falls within a certain interval \cite{meeker}.  In the case of left truncation, information about sample units are unknown before a certain time point.  Therefore, these units follow a conditional lifetime distribution where the conditioning is based on having survived up until the time they enter the sample.  Ignoring the left truncation and treating the data as identically distributed can lead to biased estimates whereas dropping left truncated observations substantially reduces the sample size.  Therefore, units exhibiting right censoring and or left truncation should be properly incorporated into the likelihood function.

\subsubsection{Likelihood}
We now define the general form for the likelihood function that takes into account left truncation, as well as right censoring.  Let $Y_{i}$ denote the lifetime of product $i$ in hours when it has failed, where $i$ goes from $i=1 \dots n$.  Let $t_i^L$ define the left truncation time in hours.  Specifically, for how many hours has a unit been running for before it is added to the sample for observation.  For units that are right censored, $c_i$ denotes the number of hours a unit has survived without failing.  We also specify the following indicator functions.  Let $\nu_i=0$ if a product is left truncated and $\nu_i=1$ otherwise.  Let $\delta_i=0$ if a product is right censored and $\delta_i=1$ otherwise.  Given the two parametric forms for $F_1$ and $F_2$ the joint likelihood for the GLFP is a function of $\bm{\theta} = (\mu_1, \sigma_1, \mu_2, \sigma_2, \pi)$.  Assuming all products are independent, the complete likelihood for one product brand is
\begin{equation*}
L(\bm{\theta})= \prod_{i=1}^{n} f(Y_i; \bm{\theta})^{\nu_i\delta_i} \left[\frac{f(Y_i;\bm{\theta})}{1-F(t_i^L;\bm{\theta})}\right]^{(1-\nu_i) \delta_i}
[1-F(c_i;\bm{\theta})]^{\nu_i (1-\delta_i)} \left[ \frac{1-F(c_i;\bm{\theta})}{1-F(t_i^L;\bm{\theta})} \right]^{(1-\nu_i)(1-\delta_i)}

A common feature of lifetime data is right censoring.  For various reasons failure times for observational units are known only to up to a lower bound. For example, observational units still surviving when the data is analyzed.  Lifetime data for products with long expected lifetimes, such as hard drives, have a natural tendency to exhibit this kind of censoring.\\

When an observation is truncated, that implies that it would not have been observed had it not coincided with the observation window. This has the existence of an observation would have been unknown if it falls within a certain interval \cite{meeker}. While right censoring restricts information to a lower bound on survival, left truncation makes information conditional on survival up to the truncation point. This is a common feature of observational data, where the factors leading to inclusion in the data set are uncontrolled. Ignoring left truncation can lead to biased estimates. However, dropping left truncated observations should be avoided because it could substantially reduce the total available information.  Therefore, we incorporate both right censoring and left truncation into the likelihood.

>>>>

@@

\subsection{Censoring and Truncation}
A common feature of lifetime data is right censoring. Due to time or budget constraints it is rare all units are observed until failure. If a unit has not yet failed when the data is analyzed it is considered right censored.  In other words, right censoring puts a lower bound on the failure time.  
Lifetime data for products with long expected lifetimes, such as hard drives, have a natural tendency to exhibit this kind of censoring.

When an observation is truncated, that implies that it would not have been observed had it not coincided with the observation window.  While right censoring restricts information to a lower bound on survival, left truncation makes information conditional on survival up to the truncation point. This is a common feature of observational data, where the factors leading to inclusion in the data set are uncontrolled. Ignoring left truncation can lead to biased estimates. However, dropping left truncated observations should be avoided because it could substantially reduce the total available information.  Therefore, we incorporate both right censoring and left truncation into the likelihood.
